Bayesian Data Analysis, Rough Draft

### 2023-12-01 Lahman Analysis/Formal Analysis/Draft1.qmd
Created Trajectory for second player

The parameter's $\beta_{n,1-4}$ describe the weights of the orthogonal polynomial for, up to the quartic, ages of players. We theorize, because Fellingham and Fisher (2017) did not explicitly articulate why, that they chose a quartic model because it has a closer fit to the data displayed in @fig-HR.By.Age.All and @fig-HR.By.Age.20to40. The properties of a squared term is that it captures the rise and fall of player's HR probability. However, the inclusion of the cubic and quartic term allow for for modeling two humps, or peaks, in a data set. This appears in @fig-HR.By.Age.20to40 where players see a peak at 27 and then their probability decreases a little, which is then followed by another peak at 31, and afterwards continues in a downward trend. This will be discussed in more depth in @sec-Model, but our model which aims to improve Fellingham and Fisher's (2017) model does not account for the cubic or quartic effects over fear of over-fitting the age data.

@@

@@

```{r}
#The selected player (id=651) is Robin Yount

set.seed(1)

select_player
```

```{r}
select_player_Batting <- Batting %>% 
  filter(playerID %in% select_player) %>% 
  select(playerID, player_index, yearID, year_index, Age, cAge, cAgeSq, AB, HR, p, park_index)

select_player_Batting
```

```{r}
#These first three effects (alpha, beta, and eta) are all corresponding to the 651st player (Robin Yount)
alpha_i651 <- fit.poly %>% 
  spread_draws(alpha[group]) %>% 
  summarise_draws() %>% 
  filter(group==651)

```{r}
#This outputs the associated means for each index for player 651
alpha_i651 %>% pull(mean)
beta_i651 %>% pull(mean)
eta_i651 %>% pull(mean)
delta_i651 %>% pull(mean)
xi_i651 %>% pull(mean)
```

```{r}
ggplot(data = select_player_Batting)+
  aes(x = cAge, y = p)+
  geom_point()+
  geom_smooth()+
  labs(title = "Actual Trajectory for player #651")
```

```{r}
ggplot(data = select_player_Batting)+
  aes(x = cAge, y=inv_logit_pred)+
  geom_point()+
  geom_smooth()+
  labs(title = "Predicted Career Trajectory for #651")
```

>>>>

The parameter's $\beta_{n,1-4}$ describe the weights of the orthogonal polynomial for, up to the quartic, ages of players. We theorize, because Fellingham and Fisher (2017) did not explicitly articulate why, that they chose a quartic model because it has a closer fit to the data displayed in @fig-HR.By.Age.All and @fig-HR.By.Age.20to40. The properties of a squared term is that it captures the rise and fall of player's HR probability. However, the inclusion of the cubic and quartic term allow for for modeling two humps, or peaks, in a data set. This appears in @fig-HR.By.Age.20to40 where players see a peak at 27 and then their probability decreases a little, which is then followed by another peak at 31, and afterwards continues in a downward trend. This will be discussed in more depth in @sec-Model but our model, which aims to improve Fellingham and Fisher's (2017) model, does not account for the cubic or quartic effects over fear of over-fitting the age data.

@@

```{r}
#Using the weights of the model, lets predict how a couple of players would do and compare it to how they actually did

#We want to set the seed of this chunk so that we're talking about the same players everytime we run the chunk
set.seed(1)

#This gives us a list of 10 players to consider so if there are a couple that look similar we can choose to not include them in the analysis
select_players <- Batting %>% 
  distinct(playerID) %>% 
  pull(playerID) %>% 
  sample(size=10, replace=FALSE)

select_players
```

```{r}
select_players_Batting <- Batting %>% 
  filter(playerID %in% select_players) %>% 
  select(playerID, player_index, yearID, year_index, Age, cAge, cAgeSq, AB, HR, p, park_index)

#select_players_Batting
```

## Robin Yount

```{r}
set.seed(1)

select_player_Batting <- Batting %>% 
  filter(playerID %in% select_player) %>% 
  select(playerID, player_index, yearID, year_index, Age, cAge, cAgeSq, AB, HR, p, park_index)
```

```{r}
# Robin Yount (651) Estimates cell

alpha_i651 <- fit.poly %>% 
  spread_draws(alpha[group]) %>% 
  summarise_draws() %>% 
  filter(group==651)

See @fig-p651Act and @fig-p651Pred.

## Pat Tabler

```{r}
# Pat Tabler (582) Estimates cell

#Find out what parks player 582 played in over his career
i582_park <- Batting %>% 
  filter(player_index == 582) %>% 
  distinct(park_index) %>% 
  pull(park_index)

#Find our what years player 582 played in over his career
i582_year <- Batting %>% 
  filter(player_index == 582) %>% 
  distinct(year_index) %>% 
  pull(year_index)

#Find the all the associated parameters needed for to predict model of player 582
#These first three are specfic estimates for 582
alpha_i582 <- fit.poly %>% 
  spread_draws(alpha[group]) %>% 
  summarise_draws() %>% 
  filter(group==582)

beta_i582 <- fit.poly %>% 
  spread_draws(beta[group]) %>% 
  summarise_draws() %>% 
  filter(group==582)

eta_i582 <- fit.poly %>% 
  spread_draws(eta[group]) %>% 
  summarise_draws() %>% 
  filter(group==582)

#Meanwhile, this value corresponds to the park
delta_i582 <- fit.poly %>% 
  spread_draws(delta[group]) %>% 
  summarise_draws() %>% 
  filter(group %in% i582_park)

#And this value corresponds to the years that he played
xi_i582 <- fit.poly %>% 
  spread_draws(xi[group]) %>% 
  summarise_draws() %>% 
  filter(group %in% i582_year)
```

```{r}
#Create the trajectory DF for player 582
i582.Trajectory <- Batting %>% 
  #Grab the player we're interested in
  filter(player_index == 582) %>% 
  #Only keep columns we're interested in
  select(playerID, player_index, yearID, year_index, Age, cAge, cAgeSq, AB, HR, p, park_index) %>%
  #Put the alpha, beta, and eta parameters in the DF
  mutate(alpha = alpha_i582 %>% pull(mean), .after = player_index) %>%
  mutate(beta = beta_i582 %>% pull(mean), .after = cAge) %>%
  mutate(eta = eta_i582 %>% pull(mean), .after = cAgeSq) %>%
  #Set-up to add "delta" to the table because 582 was traded
  right_join(delta_i582 %>% select(group, mean), by = join_by(park_index==group)) %>% 
  mutate(delta = mean, .after = park_index) %>% 
  select(!mean) %>% 
  #Set-up to add "xi" to the table because 582 was traded
  right_join(xi_i582 %>% select(group, mean), by = join_by(year_index==group)) %>% 
  mutate(xi = mean, .after = year_index) %>% 
  select(!mean) %>% 
  #Create the predictions for the player
  mutate(pred = alpha + (beta*cAge) + (eta*cAgeSq) + (delta) + (xi) , .after = p) %>% 
  mutate(inv_logit_pred = inv.logit(pred), .after=pred)
```

See @fig-p582Act and @fig-p582Pred.

```{r}

```

@@

```{r}
#| label: fig-p651Act
ggplot(data = select_player_Batting)+
  aes(x = cAge, y = p)+
  geom_point()+
  geom_smooth()+
  labs(title = "Actual Trajectory for player #651")
```

```{r}
#| label: fig-p651Pred
ggplot(data = select_player_Batting)+
  aes(x = cAge, y=inv_logit_pred)+
  geom_point()+
  geom_smooth(color="red")+
  labs(title = "Predicted Career Trajectory for #651")
```

```{r}
#| label: fig-p582Act
ggplot(data = i582.Trajectory)+
  aes(x = cAge, y = p)+
  geom_point()+
  geom_smooth()+
  labs(title = "Actual Trajectory for player #582")
```

```{r}
#| label: fig-p582Pred
ggplot(data = i582.Trajectory)+
  aes(x = cAge, y=inv_logit_pred)+
  geom_point()+
  geom_smooth(color="red")+
  labs(title = "Predicted Career Trajectory for #582")
```
